---
title: "Energy Transition Dashboard"
subtitle: "Renewable Growth and Emissions Reduction Analysis"
author: "Data Analysis Team"
date: "May 23, 2025"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: cosmo
    css: styles.css
    source_code: embed
    vertical_layout: fill
runtime: shiny
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
# Load required packages
library(flexdashboard)
library(shiny)
library(dplyr)
library(ggplot2)
library(plotly)
library(readr)
library(tidyr)
library(forcats)
library(viridis)
library(scales)
library(bslib)
library(DT)
library(ggrepel)  # Add ggrepel for better label placement

# Set overall theme for consistent plotting
theme_set(theme_minimal(base_size = 12))

# Set custom color palettes for better readability and accessibility
renewable_palette <- viridis(8, option = "D")
region_palette <- viridis(6, option = "H", begin = 0.1, end = 0.9)
emission_palette <- viridis(10, option = "B")

# Color-blind friendly palette for binary contrasts
binary_palette <- c("#0072B2", "#D55E00")

# Define constant values
NET_ZERO_YEAR <- 2050
RENEWABLE_THRESHOLD <- 60 # 60% renewable energy target
CURRENT_YEAR <- 2022 # Latest data in our dataset
EARLIEST_YEAR <- 2015 # For historical analysis

# Load the merged dataset - using explicit readr::read_csv to avoid namespace issues
merged_data <- readr::read_csv("energy_emissions_merged_2015_2022.csv")
  
# Load continent and development status aggregated data
continent_data <- readr::read_csv("continent_aggregated_2015_2022.csv")
development_data <- readr::read_csv("development_aggregated_2015_2022.csv")

# Print column names for continent data to help diagnose issues
print("Columns in continent_data:")
print(names(continent_data))
print("First few rows of continent_data:")
print(head(continent_data))

# Use continent_data directly as we already have data from 2015
continent_data_extended <- continent_data

# Prepare datasets for the research questions

# Print diagnostics to understand available columns
print("Available columns in continent_data_extended:")
print(names(continent_data_extended))

# RQ1: Renewable Share Trajectory
# Calculate growth rates and project future renewable share by region
# First check if the renewables_share column exists
if (!"renewables_share" %in% names(continent_data_extended)) {
  # Try to find an alternative column that might have the renewable share data
  possible_columns <- c("calc_renewables_share_energy", "renewables_share_energy", "renewable_share")
  for (col in possible_columns) {
    if (col %in% names(continent_data_extended)) {
      print(paste("Using", col, "as a substitute for renewables_share"))
      continent_data_extended$renewables_share <- continent_data_extended[[col]]
      break
    }
  }
}

rq1_data <- continent_data_extended %>%
  filter(!is.na(renewables_share)) %>%
  arrange(continent, year)

# Print diagnostic info about rq1_data
print("Structure of rq1_data:")
print(str(rq1_data))

# Calculate compound annual growth rates (CAGR) for renewable share
rq1_growth_rates <- rq1_data %>%
  group_by(continent) %>%
  summarize(
    min_year = min(year),
    max_year = max(year),
    start_value = renewables_share[year == min(year)],
    end_value = renewables_share[year == max(year)],
    years_diff = max_year - min_year,
    # Compound Annual Growth Rate (CAGR) formula: (end/start)^(1/years) - 1
    renewable_share_cagr = (end_value/start_value)^(1/years_diff) - 1,
    # Simple annual percentage point change
    annual_pp_change = (end_value - start_value) / years_diff,
    .groups = "drop"
  )

# Generate projected values based on both CAGR and linear trends
projection_years <- seq(CURRENT_YEAR + 1, NET_ZERO_YEAR)

# Function to generate projections for a continent
generate_continent_projections <- function(cont_name, base_data, growth_rates) {
  growth_info <- growth_rates %>% filter(continent == cont_name)
  
  if(nrow(growth_info) == 0) {
    return(NULL)
  }
  
  last_value <- base_data %>% 
    filter(continent == cont_name, year == max(year)) %>% 
    pull(renewables_share)
  
  if(length(last_value) == 0) {
    return(NULL)
  }
  
  # Generate both compound growth and linear projections
  tibble(
    continent = cont_name,
    year = projection_years,
    # CAGR-based projection
    cagr_projection = last_value * (1 + growth_info$renewable_share_cagr)^(year - CURRENT_YEAR),
    # Linear projection
    linear_projection = last_value + growth_info$annual_pp_change * (year - CURRENT_YEAR),
    # Use linear by default (tends to be more conservative for exponential trends)
    renewables_share = linear_projection,
    projection_type = "linear"
  )
}

# Create projections for each continent
continent_projections <- bind_rows(
  lapply(unique(rq1_growth_rates$continent), function(c) {
    generate_continent_projections(c, rq1_data, rq1_growth_rates)
  })
) %>%
  # Ensure projection_type is not NULL
  mutate(projection_type = ifelse(is.na(projection_type), "linear", projection_type))

# Combine historical data with projections for visualization
# First, verify that continent_projections contains the expected columns
print("Checking continent_projections columns:")
print(names(continent_projections))

# Create a safe version of the historical data
historical_data <- rq1_data %>% 
  select(continent, year, renewables_share) %>%
  mutate(projection_type = "historical")

# Create a safe version of the projection data
if (nrow(continent_projections) > 0 && "projection_type" %in% names(continent_projections)) {
  projection_data <- continent_projections %>%
    select(continent, year, renewables_share, projection_type)
} else {
  # If projection_type is missing, create it
  warning("projection_type column missing in continent_projections")
  projection_data <- if (nrow(continent_projections) > 0) {
    continent_projections %>%
      select(continent, year, renewables_share) %>%
      mutate(projection_type = "linear")
  } else {
    # If no projections available, create empty tibble with correct structure
    tibble(
      continent = character(),
      year = numeric(),
      renewables_share = numeric(),
      projection_type = character()
    )
  }
}

# Now bind them together
rq1_viz_data <- bind_rows(historical_data, projection_data) %>%
  # Ensure all rows have a projection_type
  mutate(projection_type = ifelse(is.na(projection_type), "historical", projection_type))

# Print diagnostic information about the visualization data
print("Structure of rq1_viz_data:")
print(str(rq1_viz_data))
print("Column names in rq1_viz_data:")
print(names(rq1_viz_data))
print("Unique values of projection_type:")
print(unique(rq1_viz_data$projection_type))

# Calculate the year when each continent would reach 60% renewable
# First, print diagnostic information
print("Structure of continent_projections:")
print(str(continent_projections))
print("Unique continents in continent_projections:")
print(unique(continent_projections$continent))
print("Summary of renewable share values:")
print(summary(continent_projections$renewables_share))

renewable_threshold_years <- continent_projections %>%
  group_by(continent) %>%
  filter(renewables_share >= RENEWABLE_THRESHOLD) %>%
  summarize(
    threshold_year = min(year),
    .groups = "drop"
  )

# More diagnostics for the result
print("Structure of renewable_threshold_years:")
print(str(renewable_threshold_years))

# RQ2: Renewable Growth vs Emission Decline
# Identify the 10 largest emitters in the latest year
latest_year <- max(merged_data$year, na.rm = TRUE)

top_emitters <- merged_data %>%
  filter(year == latest_year, !is.na(emissions)) %>%
  arrange(desc(emissions)) %>%
  head(10) %>%
  pull(country)

# Extract data for the top emitters
rq2_data <- merged_data %>%
  filter(country %in% top_emitters) %>%
  select(country, year, renewables_consumption, emissions, calc_renewables_share_energy) %>%
  arrange(country, year)

# Calculate annual growth rates for renewable electricity and required emission reduction
rq2_growth_rates <- rq2_data %>%
  group_by(country) %>%
  summarize(
    start_year = min(year),
    end_year = max(year),
    start_renewables = renewables_consumption[year == min(year)],
    end_renewables = renewables_consumption[year == max(year)],
    start_emissions = emissions[year == min(year)],
    end_emissions = emissions[year == max(year)],
    
    # Annual growth rate for renewables (CAGR)
    renewable_growth_rate = (end_renewables / start_renewables)^(1/(end_year - start_year)) - 1,
    
    # Annual reduction rate for emissions (observed)
    emission_reduction_rate = 1 - (end_emissions / start_emissions)^(1/(end_year - start_year)),
    
    # Required annual reduction rate to reach net zero by 2050 from the latest year
    # Formula: 1 - (target/current)^(1/years_left)
    # For net zero, target = 0, so formula simplifies to 1 - 0^(1/years_left) = 1
    years_to_net_zero = NET_ZERO_YEAR - end_year,
    required_reduction_rate = 1 - (0.05 / end_emissions)^(1/years_to_net_zero), # assuming 95% reduction as "net zero"
    
    # Is current renewable growth sufficient?
    is_renewable_growth_sufficient = renewable_growth_rate > required_reduction_rate,
    
    .groups = "drop"
  ) %>%
  mutate(
    # Format rates as percentages for display
    renewable_growth_pct = renewable_growth_rate * 100,
    emission_reduction_pct = emission_reduction_rate * 100,
    required_reduction_pct = required_reduction_rate * 100,
    
    # Gap between current and required rates
    gap_pct = emission_reduction_pct - required_reduction_pct,
    
    # Order countries by emissions gap
    country = fct_reorder(country, gap_pct)
  )


```

# Overview {.sidebar}

This dashboard analyzes global energy transition trends, focusing on two
key research questions:

### Research Questions

**RQ1 – Renewable Share Trajectory**\
How quickly has the share of renewables in primary energy grown since
2015, and which regions are on track to cross the 60% renewables
threshold before 2050?

**RQ2: Renewable Growth vs Emission Decline**\
For the ten largest emitters, does the average annual growth rate of
renewable electricity generation (2015-2022) exceed the annual
GHG-reduction rate required to hit Net-Zero by 2050?

```{r sidebar_controls}
# Empty sidebar controls section - region selection moved to Regional Renewable Share Trends
```

**Data Sources:**\
- Our World in Data (OWID) - Energy consumption\
- Climate Watch - Emissions data

**Time Range:** 2015-2022 (historical data) with projections to 2050

### Key Findings

-   **Renewable Growth:** Europe and South America show the fastest
    renewable energy adoption.\
-   **Net-Zero Gap:** All large emitters need to accelerate their
    renewable transition to meet 2050 targets. None are currently on
    track.\
-   **Regional Disparities:** Significant variation exists across
    regions in renewable energy growth rates.

# Renewable Share Trajectory {data-icon="fa-chart-line"}

## Row {.tabset data-height="600"}

### Regional Renewable Share Trends

```{r regional_controls}
# Region selection moved from sidebar to be directly with the regional trends plot
selectInput(
  "region_select", 
  "Select Region for Detailed View:", 
  choices = c("All Regions", unique(as.character(rq1_viz_data$continent))),
  selected = "All Regions"
)
```

```{r regional_trends}
renderPlotly({
  # Make a copy of rq1_viz_data to avoid modifying the original
  local_viz_data <- rq1_viz_data
  
  # Diagnostic: Check column names and data structure
  print("Original column names in rq1_viz_data:")
  print(names(rq1_viz_data))
  print("First few rows of rq1_viz_data:")
  print(head(rq1_viz_data))
  
  # Ensure projection_type column exists
  if (!"projection_type" %in% names(local_viz_data)) {
    message("Adding missing projection_type column to local_viz_data")
    local_viz_data <- local_viz_data %>% 
      mutate(projection_type = factor("historical", levels = c("historical", "linear")))
  } else {
    local_viz_data <- local_viz_data %>%
      mutate(projection_type = factor(projection_type, levels = c("historical", "linear")))
  }
  
  # Diagnostic: Verify projection_type
  print("Unique values of projection_type in local_viz_data:")
  print(unique(local_viz_data$projection_type))
  
  # Filter based on region selection
  filtered_data <- if (input$region_select == "All Regions") {
    local_viz_data
  } else {
    local_viz_data %>% filter(continent == input$region_select)
  }
  
  # Diagnostic: Check filtered_data
  print("Column names in filtered_data:")
  print(names(filtered_data))
  print("First few rows of filtered_data:")
  print(head(filtered_data))
  print("Unique values of projection_type in filtered_data:")
  print(unique(filtered_data$projection_type))
  
  # Ensure projection_type exists in filtered_data
  if (!"projection_type" %in% names(filtered_data)) {
    message("Adding missing projection_type column to filtered_data")
    filtered_data <- filtered_data %>% 
      mutate(projection_type = factor("historical", levels = c("historical", "linear")))
  } else {
    filtered_data <- filtered_data %>%
      mutate(projection_type = factor(projection_type, levels = c("historical", "linear")))
  }
  
  # Filter out NA values
  filtered_data <- filtered_data %>%
    filter(!is.na(year), !is.na(renewables_share), !is.na(continent), !is.na(projection_type))
  
  # Diagnostic: Check filtered_data after NA removal
  print("Rows in filtered_data after NA removal:")
  print(nrow(filtered_data))
  print("Unique projection_type values after NA removal:")
  print(unique(filtered_data$projection_type))
  
  # Check for valid data
  if (nrow(filtered_data) == 0 || all(is.na(filtered_data$renewables_share))) {
    p <- ggplot() +
      annotate("text", x = 0.5, y = 0.5, label = "No data available for selected region", size = 5) +
      theme_void()
    return(ggplotly(p))
  }
  
  # Create the plot with geom_text instead of geom_label_repel
  p <- ggplot(filtered_data, aes(x = year, y = renewables_share, color = continent)) +
    geom_line(aes(linetype = projection_type), linewidth = 1.2) +
    geom_hline(yintercept = RENEWABLE_THRESHOLD, linetype = "dashed", color = "gray60") +
    annotate("text", x = 2025, y = RENEWABLE_THRESHOLD + 2, 
             label = "60% Renewable Target", color = "gray40") +
    geom_point(data = renewable_threshold_years %>% 
                 inner_join(continent_projections, by = "continent") %>%
                 filter(year == threshold_year),
               aes(x = threshold_year, y = RENEWABLE_THRESHOLD),
               size = 3, shape = 21, fill = "white") +
    # Use geom_text with manual positioning to avoid overlap
    geom_text(data = renewable_threshold_years %>% 
                filter(!is.na(threshold_year)),
              aes(x = threshold_year, y = RENEWABLE_THRESHOLD + 5, 
                  label = sprintf("%s: %d", continent, threshold_year)),
              size = 3, vjust = -0.5, hjust = 0.5, check_overlap = TRUE) +
    scale_x_continuous(breaks = seq(2015, 2050, by = 5)) +
    scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 10)) +
    scale_color_manual(values = region_palette) +
    scale_linetype_manual(
      values = c("historical" = "solid", "linear" = "dashed"),
      labels = c("historical" = "Historical Data", "linear" = "Projected Trend"),
      drop = FALSE
    ) +
    labs(title = "Renewable Energy Share by Region (2015-2050)",
         subtitle = "Historical trends and projections toward 60% renewable energy target",
         x = "Year", 
         y = "Renewable Share (%)",
         color = "Region",
         linetype = "Data Type") +
    theme(legend.position = "bottom",
          panel.grid.minor = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1))
  
  # Convert to plotly with simplified legend handling
  plt <- ggplotly(p, tooltip = c("continent", "year", "renewables_share", "projection_type"))
  
  # Diagnostic: Inspect Plotly object structure
  print("Plotly object structure before layout:")
  print(str(plt$x$data, max.level = 2))
    # Manually adjust legend and improve display
  plt <- plt %>% 
    layout(
      legend = list(
        title = list(text = "Region / Data Type"),
        orientation = "h",
        y = -0.2,
        x = 0.5,
        xanchor = "center"
      ),
      margin = list(b = 150, l = 60, r = 40, t = 80),
      autosize = TRUE,
      height = 500
    )
  
  plt
})
```

### Regional Growth Rates

```{r growth_rates}
renderPlotly({    p <- ggplot(rq1_growth_rates, aes(x = fct_reorder(continent, annual_pp_change), 
                                  y = annual_pp_change, 
                                  fill = continent)) +
    geom_bar(stat = "identity", width = 0.7) +
    geom_text(aes(label = sprintf("%.2f pp/year", annual_pp_change)), 
              hjust = -0.1, size = 3.5) + # size is still valid for text
    coord_flip() +
    scale_fill_manual(values = region_palette) +
    labs(title = "Annual Renewable Share Growth by Region",
         subtitle = "Average percentage point increase per year",
         x = NULL, 
         y = "Annual Percentage Point Change",
         fill = "Region") +
    theme(legend.position = "none",
          panel.grid.major.y = element_blank())
    # Convert to plotly with improved layout for better display
  ggplotly(p, tooltip = c("continent", "annual_pp_change")) %>%
    layout(
      margin = list(r = 150, l = 40, b = 50, t = 80),
      autosize = TRUE,
      height = 450
    )
})
```

### Threshold Achievement Table

```{r threshold_table}
renderDataTable({
  # Combine growth rate and threshold year data
  threshold_table <- rq1_growth_rates %>%
    left_join(renewable_threshold_years, by = "continent") %>%
    mutate(
      current_share = end_value,
      years_to_threshold = ifelse(is.na(threshold_year), NA, threshold_year - max_year),
      status = case_when(
        is.na(threshold_year) ~ "Not on track by 2050",
        threshold_year <= 2030 ~ "On track (<2030)",
        threshold_year <= 2040 ~ "Moderate progress (2030-2040)",
        threshold_year <= 2050 ~ "Slow progress (2040-2050)",
        TRUE ~ "Not on track by 2050"
      )
    ) %>%
    select(
      Continent = continent, 
      `Current Share (%)` = current_share,
      `Annual Growth (pp)` = annual_pp_change,
      `Projected Achievement Year` = threshold_year,
      `Years to Target` = years_to_threshold,
      Status = status    )
  
  # Format the table for display
  datatable(
    threshold_table,
    options = list(
      pageLength = 6,
      dom = 'ft',
      ordering = TRUE
    ),
    rownames = FALSE
  ) %>%
    formatRound(columns = c("Current Share (%)", "Annual Growth (pp)"), digits = 2) %>%
    formatStyle(
      'Status',
      backgroundColor = styleEqual(
        c("On track (<2030)", "Moderate progress (2030-2040)", 
          "Slow progress (2040-2050)", "Not on track by 2050"),
        c('#a1d99b', '#fee8c8', '#fdbb84', '#e34a33')
      )
    )
})
```

## Row {data-height="300"}

### Historical Context

The global shift toward renewable energy has accelerated considerably
since 2015, with regional variations reflecting different energy
policies, resource availability, and economic factors. Europe has
consistently led in renewable energy adoption, driven by strong policy
frameworks like the European Green Deal. South America benefits from
abundant hydroelectric resources, contributing to its high renewable
share.

Projections based on historical growth rates show which regions are
likely to achieve the 60% renewable threshold by 2050 - a critical
benchmark for meeting the Paris Agreement goals and limiting global
warming to 1.5°C. Notably, regions with already substantial renewable
penetration (like South America) or strong policy frameworks (like
Europe) show faster transition trajectories.

### Methodology Notes

This analysis uses historical data from 2015-2022 and projects renewable
energy share based on observed growth rates. Two projection methods were
considered:

1.  **Compound Annual Growth Rate (CAGR)** - Accounts for exponential
    growth patterns
2.  **Linear Annual Growth** - Projects based on average percentage
    point increase per year

The dashboard presents linear projections as the default view as they
tend to provide more conservative estimates for policy planning
purposes. The 60% renewable threshold was selected based on the
International Energy Agency (IEA) and IPCC recommendations for energy
system transformation compatible with 1.5°C pathways.

# Emissions Gap Analysis {data-icon="fa-industry"}

## Row

### Top Emitters' Renewable Growth vs Required Reduction {.tabset}

#### Chart

```{r emissions_gap}
renderPlotly({  
  # Sort the data by emissions to match the order in the table
  sorted_data <- rq2_growth_rates %>%
    arrange(desc(end_emissions))
  
  p <- ggplot(sorted_data, aes(y = reorder(country, end_emissions))) +    # Current emission reduction rate
    geom_segment(aes(x = 0, xend = emission_reduction_pct, yend = country, 
                    color = "Current Reduction Rate"),
                linewidth = 5, alpha = 0.7) +    # Required emission reduction rate - using solid vertical lines with clear markers
    geom_segment(aes(x = required_reduction_pct, xend = required_reduction_pct,
                    y = as.numeric(reorder(country, end_emissions)) - 0.4, 
                    yend = as.numeric(reorder(country, end_emissions)) + 0.4,
                    color = "Required for Net-Zero"),
                linewidth = 1.5, alpha = 0.9) +
    # Add arrow markers to highlight required reduction targets
    geom_point(aes(x = required_reduction_pct, y = reorder(country, end_emissions), 
                  color = "Required for Net-Zero"),
              size = 4, shape = 18) + # Using diamond shape for better visibility
    # Renewable growth rate marker 
    geom_point(aes(x = renewable_growth_pct, y = reorder(country, end_emissions), 
                  color = "Renewable Growth Rate"),
              size = 5, shape = 16) + # Using circle shape for visibility
    # Add vertical lines for renewable growth rate for better comparison
    geom_segment(aes(x = renewable_growth_pct, xend = renewable_growth_pct, 
                   y = as.numeric(reorder(country, end_emissions)) - 0.3, 
                   yend = as.numeric(reorder(country, end_emissions)) + 0.3,
                   color = "Renewable Growth Rate"),
                linewidth = 1, alpha = 0.8) +
    
    # Gap indicators - clearer formatting
    geom_text(aes(x = pmax(emission_reduction_pct, required_reduction_pct) + 1,
                y = reorder(country, end_emissions),
                label = sprintf("Emission Gap: %.1f%%", gap_pct),
                color = if_else(gap_pct >= 0, "Sufficient", "Insufficient")),
             hjust = 0, size = 3.5, fontface = "bold") +
    # Clearer annotation comparing renewable growth with required reduction
    geom_text(aes(x = renewable_growth_pct + 0.5,
                y = as.numeric(reorder(country, end_emissions)) + 0.25,
                label = sprintf("Renewable Growth: %.1f%% (%.1f%% of required)", 
                              renewable_growth_pct, 
                              ifelse(required_reduction_pct > 0, renewable_growth_pct/required_reduction_pct*100, 0)),
                color = if_else(renewable_growth_pct >= required_reduction_pct, "Sufficient", "Insufficient")),
             hjust = 0, size = 2.7) +    
    scale_color_manual(values = c("Current Reduction Rate" = "#3182bd", 
                                 "Required for Net-Zero" = "#E65100",  # Changed to orange for better visibility
                                 "Renewable Growth Rate" = "#31a354",
                                 "Sufficient" = "#31a354",
                                 "Insufficient" = "#de2d26"),
                      name = "Metrics") +
    scale_x_continuous(limits = c(-5, 40)) +
    labs(title = "Emission Reduction Gap for Top 10 Emitters",
         subtitle = "Current vs Required Annual Reduction Rates (%) - All 10 Top Emitters Not on Track",
         x = "Annual Rate (%) - Negative values indicate increasing emissions",
         y = NULL) +
    theme(legend.position = "bottom",
          panel.grid.minor = element_blank())
  
  ggplotly(p) %>%
    layout(margin = list(l = 120, r = 120))  # Add margin for labels
})
```

#### Explanation

```{r chart_explanation, echo=FALSE}
HTML('
<div style="background-color: #f8f9fa; padding: 15px; border-radius: 5px; margin-bottom: 20px;">
  <p>This chart compares three key metrics for the top 10 global emitters:</p>
  <ul>
    <li><b style="color: #3182bd;">Blue bars</b> show each country\'s current annual emission reduction rate (negative values indicate increasing emissions)</li>
    <li><b style="color: #E65100;">Orange vertical lines and diamonds</b> show the required annual reduction rate to reach net-zero by 2050</li>
    <li><b style="color: #31a354;">Green circles</b> show each country\'s renewable energy growth rate</li>
    <li><b>Emission Gap</b> shows the difference between current and required reduction rates (all are negative, indicating no country is on track)</li>
    <li><b>Renewable Growth</b> shows both the percentage growth and what percentage of the required reduction rate this represents</li>
  </ul>
  <p><b>Note:</b> The comparison between renewable growth and required reduction helps assess whether countries\' clean energy transition is proceeding fast enough to support their net-zero goals.</p>
</div>
')
```

## Row {.tabset}

### Detailed Metrics Table

```{r metrics_table}
renderDataTable({
  # Order by emissions to match the emissions gap chart
  ordered_data <- rq2_growth_rates %>%
    arrange(desc(end_emissions))
    
  metrics_table <- ordered_data %>%
    select(
      Country = country,
      `Latest Emissions (Mt)` = end_emissions,
      `Renewable Growth Rate (%)` = renewable_growth_pct,
      `Current Emission Reduction (%)` = emission_reduction_pct,
      `Required Reduction Rate (%)` = required_reduction_pct,
      `Gap (% points)` = gap_pct,
      `On Track?` = is_renewable_growth_sufficient
    )
  
  datatable(
    metrics_table,
    options = list(
      pageLength = 10,
      dom = 'ft',
      ordering = TRUE
    ),
    rownames = FALSE
  ) %>%
    formatRound(columns = c("Latest Emissions (Mt)"), digits = 0) %>%
    formatRound(columns = c("Renewable Growth Rate (%)", 
                           "Current Emission Reduction (%)",
                           "Required Reduction Rate (%)",
                           "Gap (% points)"), digits = 2) %>%
    formatStyle(
      'Gap (% points)',
      backgroundColor = styleInterval(
        c(0),
        c('#fb6a4a', '#a1d99b')
      ),
      color = styleInterval(
        c(0),
        c('white', 'black')
      )
    ) %>%
    formatStyle(
      'On Track?',
      backgroundColor = styleEqual(
        c(TRUE, FALSE),
        c('#a1d99b', '#fb6a4a')
      )
    )
})
```

### Renewable Adoption Trajectories

```{r renewable_trajectories}
renderPlotly({
  # Get ordered countries based on emissions (matching main emissions gap chart)
  ordered_countries <- rq2_growth_rates %>%
    arrange(desc(end_emissions)) %>%
    pull(country)
  
  # Create time series with consistent country ordering
  ordered_data <- rq2_data %>%
    mutate(country = factor(country, levels = ordered_countries))
  
  # Create time series for the top emitters
  p <- ggplot(ordered_data, aes(x = year, y = calc_renewables_share_energy, 
                          color = country, group = country)) +
    geom_line(linewidth = 1.2) +
    geom_point(size = 3) +
    scale_color_viridis_d(option = "D") +
    labs(title = "Renewable Energy Share Trends for Top Emitters",
         subtitle = "Percentage of primary energy from renewable sources",
         x = "Year",
         y = "Renewable Share (%)",
         color = "Country") +
    theme(legend.position = "right")
  
  ggplotly(p, tooltip = c("country", "year", "calc_renewables_share_energy")) %>%
    layout(legend = list(x = 1.05, y = 0.5))
})
```

### Net-Zero Alignment

```{r netzero_alignment}
renderPlotly({
  # Get ordered countries based on emissions (matching main emissions gap chart)
  ordered_countries <- rq2_growth_rates %>%
    arrange(desc(end_emissions)) %>%
    pull(country)
  
  alignment_data <- rq2_growth_rates %>%
    # Ensure countries are ordered by emissions
    mutate(
      country = factor(country, levels = ordered_countries),
      alignment_cat = case_when(
        gap_pct >= 5 ~ "Strong alignment",
        gap_pct >= 0 ~ "Moderate alignment",
        gap_pct >= -5 ~ "Slight misalignment",
        TRUE ~ "Significant misalignment"
      ),
      alignment_cat = factor(alignment_cat, 
                            levels = c("Strong alignment", "Moderate alignment", 
                                       "Slight misalignment", "Significant misalignment"))
    )
  
  p <- ggplot(alignment_data, 
             aes(x = renewable_growth_pct, y = required_reduction_pct, 
                 color = alignment_cat, label = country)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray60") +
    geom_point(size = 5, alpha = 0.8) +
    geom_text(hjust = -0.1, vjust = 1, check_overlap = TRUE) +
    scale_color_manual(values = c("Strong alignment" = "#1a9850", 
                                 "Moderate alignment" = "#91cf60",
                                 "Slight misalignment" = "#fc8d59", 
                                 "Significant misalignment" = "#d73027")) +
    labs(title = "Net-Zero Alignment Matrix",
         subtitle = "Renewable growth rate vs. required emission reduction rate",
         x = "Renewable Growth Rate (%/year)",
         y = "Required Emission Reduction Rate (%/year)",
         color = "Alignment Category") +
    theme(legend.position = "bottom")
  
  ggplotly(p, tooltip = c("country", "renewable_growth_pct", "required_reduction_pct", "gap_pct"))
})
```

## Row {data-height="250"}

### Key Insights

The analysis reveals a significant gap between current performance and
Net-Zero requirements for major emitters. Key findings include:

-   None of the 10 largest emitters are currently on track to meet
    Net-Zero by 2050 based on their renewable energy growth rates.
-   While most countries are increasing renewable capacity, the pace
    generally falls short of what's needed for deep decarbonization.
-   Industrialized nations like Germany and the UK show relatively
    better alignment between renewable growth and emission reduction
    targets compared to other major emitters, though they still face
    significant gaps to Net-Zero.
-   Several major Asian economies show rapid renewable deployment
    growth, particularly Indonesia (20.98% annual renewable growth), but
    from low baselines, creating a substantial gap with Net-Zero
    requirements.

### Methodological Approach

This analysis compares three key metrics for each top emitter:

1.  **Renewable Energy Growth Rate** - Annual percentage growth in
    renewable electricity generation (2015-2022)
2.  **Current Emission Reduction Rate** - Observed annual percentage
    reduction in GHG emissions (2015-2022)
3.  **Required Reduction Rate** - Annual percentage reduction needed
    from current levels to reach Net-Zero by 2050

The gap between current and required rates indicates the additional
effort needed. A country is considered "on track" when its renewable
growth rate exceeds the required emission reduction rate, suggesting its
clean energy transition is proceeding at sufficient speed to support
Net-Zero ambitions. Based on this methodology, our analysis shows that
all major emitters currently have negative gaps, indicating significant
additional efforts are needed across all countries to meet Net-Zero
targets.

# Data Explorer {data-icon="fa-table"}

## Row {.tabset data-height="600"}

### Interactive Data Explorer

```{r data_explorer}
# Create a filtered dataset for exploration
explorer_data <- merged_data %>%
  select(
    Country = country,
    Year = year,
    Continent = continent,
    `Development Status` = development_status,
    `ISO Code` = iso_code,
    Population = population,
    GDP = gdp,
    `Primary Energy (TWh)` = primary_energy_consumption,
    `Fossil Fuel (TWh)` = fossil_fuel_consumption,
    `Renewables (TWh)` = renewables_consumption,
    `Emissions (Mt)` = emissions,
    `Renewable Share (%)` = calc_renewables_share_energy,
    `Fossil Share (%)` = calc_fossil_share_energy
  )

renderDataTable({
  datatable(
    explorer_data,
    extensions = 'Buttons',
    options = list(
      pageLength = 15,
      scrollX = TRUE,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel'),
      ordering = TRUE,
      searching = TRUE
    ),
    filter = 'top',
    rownames = FALSE
  ) %>%
    formatRound(columns = c("Population", "GDP", "Primary Energy (TWh)", 
                           "Fossil Fuel (TWh)", "Renewables (TWh)", "Emissions (Mt)"), 
               digits = 2) %>%
    formatRound(columns = c("Renewable Share (%)", "Fossil Share (%)"), 
               digits = 1)
})
```

## Row {data-height="400"}

### Methodology and Data Sources

**Data Sources:**\
- **Energy Data**: Our World in Data (OWID) energy dataset, which
compiles data from BP Statistical Review of World Energy and other
sources\
- **Emissions Data**: Climate Watch, which provides harmonized GHG
emissions data by country and sector

**Key Calculations:**\
1. **Renewable Share**: Percentage of primary energy consumption from
renewable sources (hydropower, solar, wind, geothermal, biomass)\
2. **Growth Rates**: Compound Annual Growth Rate (CAGR) for renewable
share and emission reductions\
3. **Net-Zero Alignment**: Comparison of current trends with required
trajectories to reach near-zero emissions by 2050

**Limitations:**\
- Historical data availability varies by country and region, with some
developing nations having less complete records\
- Projections assume continuation of current policy environments and
technological trends\
- The 60% renewable threshold is based on global models; optimal
renewable share may vary by region depending on energy needs and
available resources

For detailed methodology, please refer to the accompanying research
paper.

```{css, echo=FALSE}
/* Add custom CSS for mobile-responsive design and dark mode support */
@media (max-width: 768px) {
  .chart-stage {
    overflow-x: auto;
  }
  
  .chart-title {
    font-size: 16px;
  }
  
  .chart-subtitle {
    font-size: 12px;
  }
}

/* Vietnam case study specific styles */
.metric-box {
  padding: 15px;
  border-radius: 5px;
  background-color: rgba(255, 255, 255, 0.9);
  box-shadow: 0 2px 5px rgba(0,0,0,0.1);
  width: 22%;
  transition: transform 0.2s ease, box-shadow 0.2s ease;
}

.metric-box:hover {
  transform: translateY(-3px);
  box-shadow: 0 4px 8px rgba(0,0,0,0.15);
}


```

```{js, echo=FALSE}
// JavaScript for handling mobile display
$(document).ready(function() {
  // Adjust plot sizes for mobile screens
  function adjustForMobile() {
    if(window.innerWidth < 768) {
      $('.chart-stage').css('height', 'auto');
    }
  }
  
  // Run on page load and resize
  adjustForMobile();
  $(window).resize(adjustForMobile);
});
```

# Vietnam Case Study {data-icon="fa-flag"}

```{r vietnam_data_prep}
# Extract Vietnam data
vietnam_data <- merged_data %>%
  filter(country == "Vietnam")

# Calculate annual growth rates for Vietnam
vietnam_growth <- vietnam_data %>%
  arrange(year) %>%
  mutate(
    renewable_yoy_change = (renewables_consumption / lag(renewables_consumption) - 1) * 100,
    renewable_share_yoy_change = calc_renewables_share_energy - lag(calc_renewables_share_energy),
    emissions_yoy_change = (emissions / lag(emissions) - 1) * 100
  )

# Project Vietnam's renewable share to 2050
vietnam_share_history <- vietnam_data %>%
  select(year, renewables_share = calc_renewables_share_energy)

# Calculate average annual percentage point change in renewable share
vietnam_avg_change <- mean(vietnam_growth$renewable_share_yoy_change, na.rm = TRUE)

# Project to 2050
last_year <- max(vietnam_data$year)
last_share <- vietnam_data$calc_renewables_share_energy[vietnam_data$year == last_year]

vietnam_projection <- tibble(
  year = (last_year + 1):2050,
  renewables_share = last_share + vietnam_avg_change * (1:length((last_year + 1):2050))
)

# Combine historical and projected data
vietnam_share_all <- bind_rows(
  vietnam_share_history %>% mutate(type = "Historical"),
  vietnam_projection %>% mutate(type = "Projected")
)

# Calculate when Vietnam might reach 60% threshold
vietnam_threshold_year <- vietnam_projection %>%
  filter(renewables_share >= 60) %>%
  summarize(year = min(year)) %>%
  pull(year)

if(length(vietnam_threshold_year) == 0) {
  vietnam_threshold_message <- "Not projected to reach 60% by 2050"
} else {
  vietnam_threshold_message <- paste0("Projected to reach 60% by ", vietnam_threshold_year)
}

# Calculate emission reduction needed for Vietnam
vietnam_latest <- vietnam_data %>%
  filter(year == max(year))

vietnam_emissions_latest <- vietnam_latest$emissions
years_to_2050 <- 2050 - max(vietnam_data$year)

# Required annual reduction rate to reach net-zero by 2050
vietnam_required_reduction <- (1 - (0.05/vietnam_emissions_latest)^(1/years_to_2050)) * 100

# Current renewable growth rate
vietnam_renewable_growth <- mean(tail(vietnam_growth$renewable_yoy_change, 3), na.rm = TRUE)

# Current emission change rate
vietnam_emission_change <- mean(tail(vietnam_growth$emissions_yoy_change, 3), na.rm = TRUE)

# Gap between current and required
vietnam_gap <- vietnam_emission_change + vietnam_required_reduction
```

## Row {.tabset}

### Vietnam's Renewable Transition

```{r vietnam_renewable_plot}
renderPlotly({  p <- ggplot(vietnam_share_all, aes(x = year, y = renewables_share, color = type, group = 1)) +
    geom_line(linewidth = 1.5) +
    geom_point(data = filter(vietnam_share_all, type == "Historical"), size = 3) +
    geom_hline(yintercept = 60, linetype = "dashed", color = "darkgrey") +
    annotate("text", x = 2030, y = 62, label = "60% Renewable Target", color = "darkgrey") +
    scale_color_manual(values = c("Historical" = "#1976D2", "Projected" = "#FF7043")) +
    scale_x_continuous(breaks = seq(2015, 2050, by = 5)) +
    labs(title = "Vietnam's Renewable Energy Share Trajectory",
         subtitle = paste0("Historical data (2015-2022) and projection to 2050 - ", vietnam_threshold_message),
         x = "Year",
         y = "Renewable Share (%)",
         color = "Data Type") +
    theme_minimal() +
    theme(legend.position = "bottom")
  
  ggplotly(p) %>%
    layout(
      hovermode = "x unified",
      autosize = TRUE,
      height = 450
    )
})
```

### Emission Trends and Challenges

```{r vietnam_emissions_plot}
renderPlotly({  p <- ggplot(vietnam_data, aes(x = year, y = emissions)) +
    geom_line(linewidth = 1.5, color = "#E53935") +
    geom_point(size = 3, color = "#E53935") +
    geom_smooth(method = "lm", formula = 'y ~ x', se = FALSE, linetype = "dashed", color = "darkgrey") +
    scale_x_continuous(breaks = seq(2015, 2022, by = 1)) +
    labs(title = "Vietnam's GHG Emissions (2015-2022)",
         subtitle = paste0("Current trend vs. required reduction for Net-Zero (", round(vietnam_required_reduction, 1), "% annual reduction needed)"),
         x = "Year",
         y = "Emissions (Mt CO2e)") +
    theme_minimal()
  
  ggplotly(p) %>%
    layout(
      hovermode = "x unified",
      autosize = TRUE,
      height = 450
    )
})
```

### Energy Mix Evolution

```{r vietnam_energy_mix}
renderPlotly({    # Prepare energy mix data
  energy_mix <- vietnam_data %>%
    select(year, 
           Coal = calc_coal_share_energy, 
           Oil = calc_oil_share_energy, 
           Gas = calc_gas_share_energy, 
           Nuclear = calc_nuclear_share_energy,
           Hydro = calc_hydro_share_energy, 
           Solar = calc_solar_share_energy, 
           Wind = calc_wind_share_energy,
           Other = calc_other_renewable_share_energy) %>%
    pivot_longer(cols = -year, names_to = "source", values_to = "share")
    
  # Create the plot  
  p <- ggplot(energy_mix, aes(x = year, y = share, fill = source)) +
    geom_area() +
    scale_fill_viridis_d(option = "turbo") +
    labs(title = "Vietnam's Energy Mix Evolution (2015-2022)",
         subtitle = "Share of primary energy by source (%)",
         x = "Year", 
         y = "Share (%)",
         fill = "Energy Source") +
    theme_minimal() +
    theme(legend.position = "bottom")
  
  ggplotly(p) %>%
    layout(
      hovermode = "x unified",
      autosize = TRUE,
      height = 450
    )
})
```

## Row {data-height="250"}

### Vietnam's Energy Transition Story

Vietnam has experienced significant changes in its energy landscape
between 2015-2022. Starting with a relatively low renewable share of
primary energy at around 18% in 2015, the country has shown impressive
growth, reaching approximately 27.5% in 2022. This growth has been
driven particularly by rapid expansion of solar and wind capacity in
recent years.

Key observations:

1.  **Renewable growth acceleration**: Vietnam's renewable share growth
    has accelerated, especially after 2019, with significant increases
    in solar deployment.

2.  **Emission challenges**: Despite renewable growth, Vietnam's total
    emissions have continued to rise due to economic and population
    growth, creating a substantial gap with Net-Zero requirements.

3.  **Regional context**: As part of the Asia region, Vietnam follows
    similar trends, but with a faster renewable adoption rate than many
    regional peers.

4.  **Policy implications**: Vietnam's National Power Development Plan
    aims to increase the share of renewables significantly, but
    achieving the 60% threshold by 2050 will require sustained policy
    support and investment.

### Vietnam's Net-Zero Gap Analysis

```{r vietnam_gap_analysis}
renderUI({
  gap_color <- if(vietnam_gap >= 0) "#4CAF50" else "#F44336"
  
  HTML(paste0(
    "<div style='display: flex; justify-content: space-around; text-align: center;'>",
    
    "<div class='metric-box'>",
    "<h4>Current Renewable Growth</h4>",
    "<p style='font-size: 24px; font-weight: bold; color: #2196F3;'>", round(vietnam_renewable_growth, 1), "%</p>",
    "<p>Annual increase</p>",
    "</div>",
    
    "<div class='metric-box'>",
    "<h4>Current Emission Trend</h4>",
    "<p style='font-size: 24px; font-weight: bold; color: #FF9800;'>", round(vietnam_emission_change, 1), "%</p>",
    "<p>Annual change</p>",
    "</div>",
    
    "<div class='metric-box'>",
    "<h4>Required Reduction</h4>",
    "<p style='font-size: 24px; font-weight: bold; color: #F44336;'>", round(vietnam_required_reduction, 1), "%</p>",
    "<p>For Net-Zero by 2050</p>",
    "</div>",
    
    "<div class='metric-box'>",
    "<h4>Net-Zero Gap</h4>",
    "<p style='font-size: 24px; font-weight: bold; color: ", gap_color, ";'>", round(vietnam_gap, 1), "%</p>",
    "<p>", ifelse(vietnam_gap >= 0, "On track", "Additional reduction needed"), "</p>",
    "</div>",
    
    "</div>"
  ))
})
```
